---
title: "Untitled"
author: "Brendan F. Miller"
date: "1/16/2023"
output: html_document
---

# co_occurance

```{r}

meta <- read.csv2(file = "/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/squidpy/compute_co_occurance_data.meta.csv", row.names = 1, sep=",")

```

```{r}

meta

```

```{r}

meta <- meta %>% 
  dplyr::mutate_at(vars(x, y), as.numeric)

```

```{r}

subsets <- readRDS("data/squidpy/sp.cooccuranceData.subsets.near.subdist100.rds")
subsets200 <- readRDS("data/squidpy/sp.cooccuranceData.subsets.near.subdist200.rds")

```

```{r}

getSubsetComs <- function(com, pos, subset_list, subsetIDs, neighIDs){
  
  ## get vector to append cell annotations of interest
  annots_temp <- rep(NA, length(rownames(pos)))
  names(annots_temp) <- rownames(pos)
  
  ## append the neighbor cells
  if(!is.na(neighIDs[1])){
    for(neigh_id in neighIDs){
      annots_temp[com == neigh_id] <- neigh_id
    }
  }
  
  ## get cell ids that are part of subset
  ## these added after the neigbors in case
  ## you want to plot all CD4 T cells first
  # then color the ones that are a subset
  ## note that the order will be important
  ## because labels are overwritten in this way
  if(!is.na(subsetIDs[1])){
    for(subsetID in subsetIDs){
      cells_temp <- as.numeric(subset_list[[subsetID]])
      ## append the subset label
      annots_temp[cells_temp] <- subsetID
    }
  }
  
  annots_temp <- as.factor(annots_temp)
  return(annots_temp)
}

```

```{r}

pos <- meta[,c("x", "y")]
com <- as.factor(meta$cluster)
names(com) <- rownames(pos)


# ----------------------------------------
# Podoplanin subsets
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets,
                             neighIDs = c("T cells", "apoptotic tumor cell"),
                             subsetIDs = c("T cells_near_apoptotic tumor cell"))
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

# ggplot2::ggsave(filename ="sim_circSubset.v2.AnearB.pdf",
#                 path = figpath, device = "pdf",
#                 plot = plt,
#                 dpi = 600, width = 4, height = 3, units = "in")

```

```{r}

pos <- meta[,c("x", "y")]
com <- as.factor(meta$cluster)
names(com) <- rownames(pos)


# ----------------------------------------
# Podoplanin subsets
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets200,
                             neighIDs = c("T cells", "apoptotic tumor cell"),
                             subsetIDs = c("T cells_near_apoptotic tumor cell"))
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

# ggplot2::ggsave(filename ="sim_circSubset.v2.AnearB.pdf",
#                 path = figpath, device = "pdf",
#                 plot = plt,
#                 dpi = 600, width = 4, height = 3, units = "in")

```

```{r}

dataDir <- "/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/squidpy/"

figpath <- "/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/plots/squidpy/"

```

```{r}

outfile <- paste0(dataDir, "sp.cooccuranceData.near.binom.subdist200.dist200.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$resolution %in% c(100,200,300,400,500,600,700,800),]

dat <- dat[grepl(pattern = "basal CK tumor cell_near_", x = dat$reference),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.cooccuranceData.subdist200.dist200.subsets.basalCK.pdf"),
          width = 30, height = 30)

```

```{r}

outfile <- paste0(dataDir, "sp.cooccuranceData.near.binom.subdist100.dist100.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$resolution %in% c(100,200,300,400,500,600,700,800),]

dat <- dat[grepl(pattern = "basal CK tumor cell_near_", x = dat$reference),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.cooccuranceData.subdist100.dist100.subsets.basalCK.pdf"),
          width = 30, height = 30)

```

```{r}

outfile <- paste0(dataDir, "sp.cooccuranceData.near.binom.subdist200.dist100.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$resolution %in% c(100,200,300,400,500,600,700,800),]

dat <- dat[grepl(pattern = "basal CK tumor cell_near_", x = dat$reference),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.cooccuranceData.subdist200.dist100.subsets.basalCK.pdf"),
          width = 30, height = 30)

```

```{r}

outfile <- paste0(dataDir, "sp.cooccuranceData.pairwise.results.10-300.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

# dat <- dat[dat$dist %in% c("100", "200", "300"),]
# dat <- dat[dat$resolution %in% c(50,100,200,300,500,700,800,1000,3000),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.cooccuranceData.pairwise.results.10-300.pdf"),
          width = 30, height = 30)

```

```{r}


outfile <- paste0(dataDir, "sp.cooccuranceData.pairwise.results.10-300.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$dist %in% c("200"),]
dat <- dat[dat$resolution %in% c(50,100,200,300,500,700,800,1000,3000),]

plotTrendsOverlay(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.ripleyData.pairwise.overlayed.pdf"),
          width = 4, height = 40)

```

# ripley

```{r}

meta_ripley <- read.csv2(file = "/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/squidpy/compute_ripley_data.meta.csv.gz", row.names = 1, sep=",")

```

```{r}

meta_ripley

```

```{r}

meta_ripley <- meta_ripley %>% 
  dplyr::mutate_at(vars(x, y), as.numeric)

```


```{r}

subsets100 <- readRDS("/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/squidpy/sp.ripleyData.subsets.near.subdist100.rds")
subsets200 <- readRDS("/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/squidpy/sp.ripleyData.subsets.near.subdist200.rds")
subsets300 <- readRDS("/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/squidpy/sp.ripleyData.subsets.near.subdist300.rds")

```

```{r, fig.height=18, fig.width=18}

vizEachCluster(object = meta_ripley[,c("x", "y")], clusters = as.factor(meta_ripley$cluster), s = 2)

```

```{r}

pos <- meta_ripley[,c("x", "y")]
com <- as.factor(meta_ripley$cluster)
names(com) <- rownames(pos)

# ----------------------------------------
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets100,
                             neighIDs = c("CA1_CA2_CA3_Subiculum", "Interneurons", "DentatePyramids"),
                             subsetIDs = NA)
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

# ----------------------------------------
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets100,
                             neighIDs = c("CA1_CA2_CA3_Subiculum", "Interneurons", "DentatePyramids"),
                             subsetIDs = c("CA1_CA2_CA3_Subiculum_near_DentatePyramids"))
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

# ----------------------------------------
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets200,
                             neighIDs = c("CA1_CA2_CA3_Subiculum", "Interneurons", "DentatePyramids"),
                             subsetIDs = c("CA1_CA2_CA3_Subiculum_near_DentatePyramids"))
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

# ----------------------------------------
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets300,
                             neighIDs = c("CA1_CA2_CA3_Subiculum", "Interneurons", "DentatePyramids"),
                             subsetIDs = c("CA1_CA2_CA3_Subiculum_near_DentatePyramids"))
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

# ----------------------------------------
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets100,
                             neighIDs = c("DentatePyramids"),
                             subsetIDs = c("CA1_CA2_CA3_Subiculum_near_DentatePyramids"))
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

# ----------------------------------------
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets200,
                             neighIDs = c("DentatePyramids"),
                             subsetIDs = c("CA1_CA2_CA3_Subiculum_near_DentatePyramids"))
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

# ----------------------------------------
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets300,
                             neighIDs = c("DentatePyramids"),
                             subsetIDs = c("CA1_CA2_CA3_Subiculum_near_DentatePyramids"))
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

# ----------------------------------------
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets300,
                             neighIDs = c("DentatePyramids", "CA1_CA2_CA3_Subiculum"),
                             subsetIDs = NA)
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

# ggplot2::ggsave(filename ="sim_circSubset.v2.AnearB.pdf",
#                 path = figpath, device = "pdf",
#                 plot = plt,
#                 dpi = 600, width = 4, height = 3, units = "in")

```

```{r}

outfile <- paste0(dataDir, "sp.ripleyData.near.binom.subdist100.dist100.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$resolution %in% c(50,100,200,300,500,700,800,1000,3000),]

dat <- dat[grepl(pattern = "CA1_CA2_CA3_Subiculum_near_", x = dat$reference),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.ripleyData.subdist100.dist100.subsets.CAs.pdf"),
          width = 30, height = 30)

```

```{r}

outfile <- paste0(dataDir, "sp.ripleyData.near.binom.subdist200.dist200.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$resolution %in% c(50,100,200,300,500,700,800,1000,3000),]

dat <- dat[grepl(pattern = "CA1_CA2_CA3_Subiculum_near_", x = dat$reference),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.ripleyData.subdist200.dist200.subsets.CAs.pdf"),
          width = 30, height = 30)

```

```{r}

outfile <- paste0(dataDir, "sp.ripleyData.near.binom.subdist300.dist300.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$resolution %in% c(50,100,200,300,500,700,800,1000,3000),]

dat <- dat[grepl(pattern = "CA1_CA2_CA3_Subiculum_near_", x = dat$reference),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.ripleyData.subdist300.dist300.subsets.CAs.pdf"),
          width = 30, height = 30)

```

```{r}

outfile <- paste0(dataDir, "sp.ripleyData.near.binom.subdist300.dist300.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$resolution %in% c(50,100,200,300,500,700,800,1000,3000),]

dat <- dat[grepl(pattern = "Astrocytes_near_", x = dat$reference),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.ripleyData.subdist300.dist300.subsets.Astro.pdf"),
          width = 30, height = 30)

```

```{r}

outfile <- paste0(dataDir, "sp.ripleyData.pairwise.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$dist %in% c("100", "200", "300"),]
dat <- dat[dat$resolution %in% c(50,100,200,300,500,700,800,1000,3000),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.ripleyData.pairwise.pdf"),
          width = 30, height = 30)

```

```{r}


outfile <- paste0(dataDir, "sp.ripleyData.pairwise.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$dist %in% c("200"),]
dat <- dat[dat$resolution %in% c(50,100,200,300,500,700,800,1000,3000),]

plotTrendsOverlay(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.ripleyData.pairwise.overlayed.pdf"),
          width = 4, height = 40)

```


```{r}


outfile <- paste0(dataDir, "sp.slideseqv2.pairwise.30-300.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$dist %in% c(30, 50, 100, 200, 300),]
dat <- dat[dat$resolution %in% c(30, 50, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1200, 1500, 3000, 6000),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.slideseqv2.pairwise.30-300.results.pdf"),
          width = 30, height = 30)

```

```{r}


outfile <- paste0(dataDir, "sp.slideseqv2.pairwise.30-300.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$dist %in% c(100),]
dat <- dat[dat$resolution %in% c(30, 50, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1200, 1500, 3000, 6000),]

plotTrendsOverlay(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.slideseqv2.pairwise.30-300.results.overlay.dist100.pdf"),
          width = 4, height = 30)

```


```{r}

outfile <- paste0(dataDir, "sp.slideseqv2.near.binom.subdist100.dist30-200.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$dist %in% c(30, 50, 100, 200),]
dat <- dat[dat$resolution %in% c(30, 50, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1200, 1500, 3000, 6000),]

dat <- dat[grepl(pattern = "CA1_CA2_CA3_Subiculum_near_", x = dat$reference),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.slideseqv2.subdist100.dist30-200.results.CAs.pdf"),
          width = 30, height = 30)

```

```{r}

outfile <- paste0(dataDir, "sp.slideseqv2.near.binom.subdist200.dist30-200.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$dist %in% c(30, 50, 100, 200),]
dat <- dat[dat$resolution %in% c(30, 50, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1200, 1500, 3000, 6000),]

dat <- dat[grepl(pattern = "CA1_CA2_CA3_Subiculum_near_", x = dat$reference),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.slideseqv2.subdist200.dist30-200.results.CAs.pdf"),
          width = 30, height = 30)

```

```{r}

outfile <- paste0(dataDir, "sp.slideseqv2.near.binom.subdist300.dist30-200.results.rds")
dat <- readRDS(outfile)

dat <- meltResultsList(resultsList = dat, id = NA)
colnames(dat) <- c("resolution", "neighbor", "Z", "reference", "dist", "subsetDist")

dat <- dat[dat$dist %in% c(30, 50, 100, 200),]
dat <- dat[dat$resolution %in% c(30, 50, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1200, 1500, 3000, 6000),]

dat <- dat[grepl(pattern = "CA1_CA2_CA3_Subiculum_near_", x = dat$reference),]

plotTrends(results = dat, idcol = "dist", figPath = paste0(figpath, "sp.slideseqv2.subdist300.dist30-200.results.CAs.pdf"),
          width = 30, height = 30)

```


# shuffled data for squidpy

```{r}

randomcellslist <- readRDS(file = "/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/squidpy/sp.ripleyData.shuffled_res50-3000.rds")

```

```{r}

meta_ripley$shuffled <- randomcellslist$`3000`$`1`

```


```{r, fig.height=18, fig.width=18}

vizEachCluster(object = meta_ripley[,c("x", "y")], clusters = as.factor(meta_ripley$shuffled), s = 2)

```

```{r}

m <- meta_ripley[,c("x", "y", "cluster", "shuffled")]
colnames(m) <- c("x", "y", "cluster", "shuffled")
write.csv2(x = m, file = "/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/squidpy/ripley.meta.csv", row.names = TRUE)

```

```{r}

pos <- meta_ripley[,c("x", "y")]
com <- as.factor(meta_ripley$cluster)
names(com) <- rownames(pos)

# ----------------------------------------
annots_temp <- getSubsetComs(com = com,
                             pos = pos,
                             subset_list = subsets300,
                             neighIDs = c("CA1_CA2_CA3_Subiculum", "Astrocytes"),
                             subsetIDs = c("Astrocytes_near_CA1_CA2_CA3_Subiculum"))
plt <- vizAllClusters(object = pos,
               clusters = annots_temp,
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

```

# four i

```{r}

meta_fouri <- read.csv2(file = "/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/squidpy/four_i_data.meta.csv.gz", row.names = 1, sep=",")

```

```{r}

plt <- vizAllClusters(object = meta_fouri[,c("x", "y")],
               clusters = as.factor(meta_fouri[,c("cluster")]),
               title = "id",
               axisAdj = 1, s = 0.8, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

```

```{r}

min(meta_fouri$x)
max(meta_fouri$x)

min(meta_fouri$y)
max(meta_fouri$y)

```

```{r}

table(meta_fouri$cluster)

```

```{r}

randomcellslist <- readRDS(file = "/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/squidpy/sp.fouri.shuffled_res50-800.rds")

```

```{r}

meta_fouri$shuffled <- randomcellslist$`800`$`1`

```


```{r}

plt <- vizAllClusters(object = meta_fouri[,c("x", "y")],
               clusters = as.factor(meta_fouri[,c("shuffled")]),
               title = "id",
               axisAdj = 1, s = 4, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")
plt

```

huge slowdown and memory problem with full dataset:

it's basically like every single pixel is now being considered a "cell", although in this case its labeled by regions.
But this means that my distances of "50" mean 50 pixels, but also means than the intersection with d=50 captures # cells = to the number of pixels of the area of circle with radius 50. This is a lot of cells. And for regions with like 30K pixels, this can add up to millions of "cells". No wonder the memory and speed are so bad for this.

Honestly the triplets could be faster once the subsets are made because will hopefully break down these groups more

So it's not just a matter of so many points, but that they are packed together with no gaps.



# sim circles


```{r}

## initialize circle params

locs <- list(
  c(0.5, 0.5)
  )

cts <- list(
  list(outer = "B", inner = c("C"))
  )

probs <- list(
  list(outer = c(1), inner = c(1))
  )

## background params
s <- 10000
cb <- c("A")
pb <- c(1)

# note that the simulation starts on 0-1 scale but is scaled to 0-3100 in function simulate_circles
# can treat these values like microns

## outer ring approx 0.01 or 3100um = 31 micron width
## inner approx 868 microns across (3100*0.28)
rads <- replicate(4, list(outer = 0.15, inner = 0.14), simplify=FALSE)
pos_30 <- simulate_circles(simulate_background(size = s, cts = cb, prob = pb),
                        locs, radii = rads, cts, probs)
pos_30$id <- 30

## outer ring approx 0.03 or 3100um = 93 micron width
## inner approx 744 microns across (3100*0.24)
rads <- replicate(4, list(outer = 0.15, inner = 0.12), simplify=FALSE)
pos_90 <- simulate_circles(simulate_background(size = s, cts = cb, prob = pb),
                        locs, radii = rads, cts, probs)
pos_90$id <- 90

## outer ring approx 0.05 or 3100um = 155 micron width
## inner approx 620 microns across (3100*0.20)
rads <- replicate(4, list(outer = 0.15, inner = 0.10), simplify=FALSE)
pos_150 <- simulate_circles(simulate_background(size = s, cts = cb, prob = pb),
                        locs, radii = rads, cts, probs)
pos_150$id <- 150

## outer ring approx 0.10 or 3100um = 310 micron width
## inner approx 310 microns across (3100*0.10)
rads <- replicate(4, list(outer = 0.15, inner = 0.05), simplify=FALSE)
pos_300 <- simulate_circles(simulate_background(size = s, cts = cb, prob = pb),
                        locs, radii = rads, cts, probs)
pos_300$id <- 300

## outer ring approx 0.13 or 3100um = 403 micron width
## inner approx 124 microns across (3100*0.04)
rads <- replicate(4, list(outer = 0.15, inner = 0.02), simplify=FALSE)
pos_400 <- simulate_circles(simulate_background(size = s, cts = cb, prob = pb),
                        locs, radii = rads, cts, probs)
pos_400$id <- 400


pos_1circle_sizes <- dplyr::bind_rows(list(
  pos_30,
  pos_90,
  pos_150,
  pos_300,
  pos_400
))

pos_1circle_sizes$id <- factor(pos_1circle_sizes$id, ordered = TRUE, levels = c("30", "90", "150", "300", "400"))

```

```{r}

dat <- pos_1circle_sizes[pos_1circle_sizes$id == "300",]
ggplot2::ggplot(data = dat) +
  ggplot2::geom_point(ggplot2::aes(x = x, y = y, color = type), size = 0.2) +
  ggplot2::scale_color_manual(values = c("A" = "#F8766D", "B" = "#00B81F", "C" = "#00A5FF")) +
  ggplot2::coord_equal() +
  ggplot2::guides(color = ggplot2::guide_legend(title = "cell type"))
# ggplot2::ggsave(filename ="slide3_circle.pdf", path = figpath, device = "pdf", dpi = 600, width = 4, height = 3, units = "in")


```

```{r}

dat$x <- dat$x * 3100
dat$y <- dat$y * 3100

```

```{r}

write.csv(x = dat[,c("x", "y", "type")], file = "/Users/brendan/Desktop/PostDoc/work/HuBMAP/repos/multiscale_celltype_colocalization_analysis/data/sim/simPairwise.meta.csv", row.names = TRUE, sep = ",")

```





