---
title: "CRAWDAD Tutorial"
author: 
- "Rafael dos Santos Peixoto"
- "Brendan F. Miller"
- "Jean Fan"
date: "10/04/2023"
output: html_document
vignette: >
  %\VignetteIndexEntry{CRAWDAD Vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide')
```

# Installation

```{r, eval=FALSE}

require(remotes)
remotes::install_github('JEFworks-Lab/CRAWDAD')

```

```{r}

library(crawdad)
library(tidyverse)

```

# Tutorial

```{r}

ncores <- 2

```

## Load simulated dataset

```{r}

data(sim)

## visualize
plt <- crawdad::vizAllClusters(cells = sim,
                               coms = sim$celltypes,
                               title = "sim",
                               axisAdj = 1, s = 6, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))
plt

## convert to SP
cells <- crawdad::toSF(pos = sim[,c("x", "y")],
                       cellTypes = sim$celltypes)
cells

```

Convert the data.frame of cells to an `sp::SpatialPointsDataFrame` object because CRAWDAD builds upon the `sf` library in R.

## Make shuffled background

`CRAWDAD` identifies cell type spatial relationships by comparing cell type organizational patterns in the real data to a set of null distributions, which are a datasets in which cell labels have been shuffled at different scales, or resolutions. We can generate this list of shuffled datasets with the following code:

```{r, warning=FALSE}

## generate background
shuffle.list <- crawdad::makeShuffledCells(cells,
                          scales = c(100, 200, 500, 800, 1000),
                          perms = 3,
                          ncores = ncores,
                          seed = 1,
                          verbose = FALSE)

```

## Run pairwise analysis

We can identify trends that describe spatial relationships between pairwise combinations of cell types in our data. `dist` refers to the distance at which neighbor cells are defined. In this example, we assess if the neighbors of each cell type are enriched or depleted in cells of another given cell type compared to each shuffled scale of the data.

```{r, warning=FALSE}

## find trends, passing background as parameter
results <- crawdad::findTrends(cells,
                        neighDist = 50,
                        shuffleList = shuffle.list,
                        ncores = ncores,
                        verbose = FALSE, 
                        returnMeans = FALSE)

```

```{r}

## convert results to data.frame
dat <- crawdad::meltResultsList(results, withPerms = T)

```

```{r}

## calculate the zscore for the multiple-test correction
zsig <- correctZBonferroni(dat)

```

## Summary Visualization

Summary visualization of CRAWDAD’s multi-scale cell-type spatial relationship analysis. The size of the dot represents the scale in which the relationship becomes significant and the color represents the zscore at that scale.

```{r colocalization, warning=FALSE}

vizColocDotplot(dat, reorder = TRUE, zSigThresh = zsig, zScoreLimit = zsig*2) +
  theme(legend.position='right',
        axis.text.x = element_text(angle = 45, h = 0))

```


## Visualizing trends

We can now visualize trends for different pairwise cell type combinations using the following code:

By default, `vizTrends()` is setup to facet wrap reference and neighbor cell type trends. Reference cell types are column headers and neighbor cell types are row headers.

```{r}

plt <- crawdad::vizTrends(dat = dat, withPerms = T)
plt 

```

We can also turn off the facet wrapping, and then choose to plot multiple trend lines in the same plot using the `id` column (or another column) in the melted results data.frame. This can be useful if we want to compare the overlay of two trends:

```{r}

## select different trend combinations
d1 <- dat[dat$reference == "A" & dat$neighbor == "A",]
plt <- crawdad::vizTrends(dat = d1, withPerms = T) +
  ggplot2::scale_x_log10()
plt 

d2 <- dat[dat$reference == "A" & dat$neighbor == "B",]
plt <- crawdad::vizTrends(dat = d2, withPerms = T) +
  ggplot2::scale_x_log10()
plt 

## combine the trends into one data.frame, and have the "id" column label the combo, plot both lines on same plot
## turn off the facet wrap so just coloring the two trends, which are labeled using the "id" column

d1$id <- "A vs A"
d2$id <- "A vs B"

d <- dplyr::bind_rows(list(d1, d2))

plt <- crawdad::vizTrends(dat = d, withPerms = T, facet = FALSE) +
  ggplot2::scale_x_log10()
plt 

```

## Defining subsets

We can also further subdivide cell types into subsets by looking for cell types whose neighbors are enriched in another particular cell type. We can do this by performing a binomial test to see if a cell's neighbors are significantly enriched in a particular cell type compared to the overall probability of observing that particular cell type.

For this dataset, we define a neighbor distance of 100 to characterize cells into subsets.

First, we generate a probability matrix where rows are each cell in the dataset, columns are the cell type labels, and values are the probability of each cell being enriched in neighbors of a given cell type.

```{r, warning=FALSE}

binomMat <- crawdad::binomialTestMatrix(cells,
                               neighDist = 50,
                               ncores = ncores,
                               verbose = TRUE)

head(binomMat)

```

We can now assign cells to subsets based on how significant their neighbors are enriched with a given cell type. For finding subsets of cells that are enriched with a given cell type, we can set the `subType` to "near", with a `subThresh` of 0.05 and sub-categorize cells of a given cell type if they are enriched with another cell type at a p-value of 0.05 or less. 

Conversely, we can also set `subType` to "away" and sub-categorize cells of a given cell type if they are depleted with a given cell type. In this case, setting the `subThresh` to above 0.5 would be advised, because in this way, cells above a p-value of 0.5 in terms of testing for enrichment would be selected. In other words, only the cells that couldn't even pass a p-value cutoff of 0.5 would be selected for, and these would be expected to be very much depleted or separated from the neighbor cell type they are being compared to.

In this tutorial, we will define subsets of cell types whose neighbors are enriched in another cell type, where neighbors were defined using a distance of 100 and a p-value threshold of 0.05:

```{r}

subset.list <- crawdad::selectSubsets(binomMat,
                             cells$celltypes,
                             subType = "near",
                             subThresh = 0.05,
                             ncores = ncores,
                             verbose = TRUE)

subset.list["C_near_B"]

```

For example, these cell ids are the cells of cell type "C" that are enriched with cell type "B".

## Visualize the subsets

To visualize how the subsets are defined, let's first look at just the "C" cells and the "B" cells

```{r}

plt <- crawdad::vizClusters(cells = cells,
                            ofInterest = c("B", "C")) + ## just color these cell types
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))

plt

```

Next, let's redefine the labels of cells that are part of the subset "C_near_B" using `selectLabels()`. Then, let's specifically visualize cells of cell type "C", "B", and also the cells are are "C_near_B"

```{r}
cells_sub <- cells

## make a temporary cell type annotation factor with labeling cells that are of a specific subset
annots_temp <- crawdad::selectLabels(df = cells,
                                     com = cells$celltypes,
                                     subsetList = subset.list,
                                     cellIDs = c("A", "B", "C", "D"), ## still keep labels for these cell types
                                     subsetIDs = c("C_near_B")) ## specifically label cells that are defined as this subset in `subset.list`

## finally, for clarity sake, we'll also rename the "C" cells that are not part of the "C_near_B"
annots_temp <- dplyr::recode_factor(annots_temp, "C" = "the other Cs")
cells_sub$celltypes <- annots_temp

## visualize the subset only
plt <- crawdad::vizClusters(cells = cells_sub,
                            ofInterest = c("B", "the other Cs", "C_near_B")) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))

plt

```
