---
title: "1_simulations"
author: 
- "Brendan F. Miller"
- "Rafael dos Santos Peixoto"
date: "10/02/2023"
# output: rmarkdown::html_document
output:
  md_document:
    variant: markdown_github
vignette: >
  %\VignetteIndexEntry{CRAWDAD Vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r}

library(crawdad)
library(tidyverse)

```

```{r}

ncores = 7

```

# Load data


Load data.frame of cell positions and labels and convert it to an `sf` object. This is because CRAWDAD builds upon the `sf` library in R.

```{r, warning=FALSE}

data(sim)

## convert to sf
cells <- crawdad:::toSF(pos = sim[,c("x", "y")],
                        celltypes = sim$celltypes)

```

# Make shuffled background

`CRAWDAD` identifies cell type spatial relationships by comparing cell type organizational patterns in the real data to a set of null distributions, which are a datasets that the cell labels have been shuffled at different scales, or resolutions. We can generate this list of shuffled datasets with the following code:

```{r, warning=FALSE}

## generate background
shuffle.list <- crawdad::makeShuffledCells(cells,
                          scales = seq(100, 1000, by=100),
                          perms = 3,
                          ncores = ncores,
                          seed = 1,
                          verbose = TRUE)

```

# Run pairwise analysis

We can identify trends that describe spatial relationships between pairwise combinations of cell types in our data. `dist` refers to the distance at which neighbor cells are defined. In this example, we assess if the neighbors of each cell type are enriched or depleted in cells of another given cell type compared to each shuffled scale of the data.

```{r, warning=FALSE}

## find trends, passing background as parameter
results <- crawdad::findTrends(cells,
                        dist = 100,
                        shuffle.list = shuffle.list,
                        ncores = ncores,
                        verbose = TRUE, 
                        returnMeans = FALSE)

```

```{r}

## convert results to data.frame
dat <- crawdad::meltResultsList(results, withPerms = T)

```

# Visualize results

```{r}

## calculate the zscore for the multiple-test correction
zsig <- correctZBonferroni(dat)

```

Summary visualization of CRAWDAD’s multi-scale cell-type spatial relationship analysis.

```{r colocalization, warning=FALSE}

vizColocDotplot(dat, reorder = TRUE, zsig.thresh = zsig, zscore.limit = zsig*2) +
  theme(legend.position='right',
        axis.text.x = element_text(angle = 45, h = 0))

```

Visualize specific trends.

```{r c_b}

dat_filter <- dat %>% 
  filter(reference == 'C') %>% 
  filter(neighbor == 'B')
vizTrends(dat_filter, lines = T, withPerms = T, sig.thresh = zsig)

```

```{r a_b}

dat_filter <- dat %>% 
  filter(reference == 'A') %>% 
  filter(neighbor == 'B')
vizTrends(dat_filter, lines = T, withPerms = T, sig.thresh = zsig)

```