---
title: "seqFISH Analysis"
author: "Brendan F. Miller"
date: "2/15/2023"
output: rmarkdown::html_document
# output:
#   md_document:
#     variant: markdown_github
vignette: >
  %\VignetteIndexEntry{seqFISH analysis}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

This vignette will go through analyses to reproduce the results and figures of the SeqFISH dataset.

```{r}

library(crawdad)
library(dplyr)

```

```{r}

ncores = 7

```

# Load data

```{r, eval=FALSE}

seq <- read.csv2(file = paste0(here::here(), "/data/seqfish/seqfish.meta.csv.gz"), row.names = 1, sep = ",")
seq <- seq[,c("x", "y", "cluster")]
## make sure the coordinates are numeric
seq <- seq %>%
  dplyr::mutate_at(vars(x, y), as.numeric)

## convert to SP
seq <- crawdad:::toSP(pos = seq[,c("x", "y")],
                        celltypes = seq$cluster)
seq

```

```{r}
## the above dataset has been saved as an rda file and able to load here:
data(seq)

## convert to SP
seq <- crawdad:::toSP(pos = seq[,c("x", "y")],
                        celltypes = seq$celltypes)
seq

```

# Visualize celltypes

```{r, fig.height=24, fig.width=24}

crawdad::vizEachCluster(cells = seq,
                        coms = as.factor(seq$celltypes),
                        s = 2)

```

# Make shuffled background

```{r}

resolutions = c(100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1200, 1500, 3000, 6000)

```

```{r}
## generate background
shuffle.list <- crawdad:::makeShuffledCells(seq,
                          resolutions = resolutions,
                          perms = 1,
                          ncores = ncores,
                          seed = 1,
                          verbose = TRUE)

## note: 0.77 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7
```

save shuffled object
```{r, eval=FALSE}

saveRDS(shuffle.list, paste0(here::here(), "/data/seqfish/sp.seqfish.shuffled_res100-6000.rds"))

```

# Run pairwise analysis

```{r}

## find trends, passing background as parameter
results <- crawdad::findTrends(seq,
                        dist = 100,
                        shuffle.list = shuffle.list,
                        perms = 1,
                        ncores = ncores,
                        seed = 1, 
                        verbose = TRUE)

## note: 1.02 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7

```

save pairwise results object
```{r, eval=FALSE}

saveRDS(results, paste0(here::here(), "/data/seqfish/sp.seqfish.pairwise.results.dist100.res100-6000.rds"))

## object with distances 50-300: sp.seqfish.pairwise.50-300.results.res100-6000.removeDups.rds

```

# Visualizing trends

```{r}

crawdad::plotTrends(results = results["Intermediate mesoderm"], idcol = "dist")

```

```{r}

crawdad::plotTrends(results = results["Endothelium"], idcol = "dist")

```

# Defining susbets

```{r}

binomMat <- crawdad::binomialTestMatrix(seq,
                               neigh.dist = 100,
                               ncores = ncores,
                               verbose = TRUE)

head(binomMat)

## note: 2.49 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7
## previous code iteration on rockfish took 21 minutes with 32 cores! much better now 

```

save the binomial matrix
```{r, eval=FALSE}

saveRDS(binomMat, paste0(here::here(), "/data/seqfish/sp.seqfish.binomMat.near.subdist100.rds"))

```

```{r}

subset.list <- crawdad::selectSubsets(binomMat,
                             seq$celltypes,
                             sub.type = "near",
                             sub.thresh = 0.05,
                             ncores = ncores,
                             verbose = TRUE)

```

save the subsets
```{r, eval=FALSE}

saveRDS(subset.list, paste0(here::here(), "/data/seqfish/sp.seqfish.subsets.near.subdist100.rds"))

```

# Run analysis on subsets

```{r}

results.subsets <- crawdad::findTrends(seq,
                        dist = 100,
                        shuffle.list = shuffle.list,
                        subset.list = subset.list,
                        perms = 1,
                        ncores = ncores,
                        seed = 1, 
                        verbose = TRUE)

## note: 11.68 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7; vs 7.53 minutes 14 cores on rockfish

```

save the results
```{r, eval=FALSE}

saveRDS(results.subsets, paste0(here::here(), "/data/seqfish/sp.seqfish.subset.results.dist100.res100-6000.rds"))

## sp.seqfish.triplet.near.binom.subdist100.dist100.results.res100-6000.removeDups.rds

```

# ---------------------------------
# Figures

stored on `brendan` branch in:
`/plots/seqfish/`

## load data

```{r, eval=FALSE}

## folder on `brendan` branch
figpath <- paste0(here::here(), "/plots/seqfish")

shuffle.list <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.shuffled_res100-6000.rds"))

binomMat <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.binomMat.near.subdist100.rds"))
subset.list <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.subsets.near.subdist100.rds"))

results <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.pairwise.results.dist100.res100-6000.rds"))
results.subsets <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.subset.results.dist100.res100-6000.rds"))

```

## Fig 2 - mesoderms on tissue

```{r, fig.height=4, fig.width=4}

plt <- crawdad::vizAllClusters(cells = seq,
                        coms = as.factor(seq$celltypes),
                        ofInterest = c("Lateral plate mesoderm", "Intermediate mesoderm"),
                        s = 2) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))
  ggplot2::theme(legend.position="none")

ggplot2::ggsave(filename = "2A_mesoderm_tissue.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))
plt

```

### zoom in

```{r}

## shuffling grid
grid <- sf::st_make_grid(seq, cellsize = 1000)
grid_coords <- as.data.frame(sf::st_coordinates(grid))

## centers of the grids to add the tile IDs
grid_coords_centroids <- as.data.frame(sf::st_coordinates(sf::st_centroid(grid)))
grid_coords_centroids$name <- as.character(rownames(grid_coords_centroids))


plt <- crawdad::vizAllClusters(cells = seq,
                               coms = seq$celltypes,
                               ofInterest = c("Lateral plate mesoderm", "Intermediate mesoderm"),
                               title = "seq",
                               axisAdj = 1, s = 6, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  
  ## add in the grid information on top of the plot
  ggplot2::geom_sf(data = grid, fill = NA) +
  ggplot2::geom_text(data = grid_coords_centroids, ggplot2::aes(X, Y, label = name))
  
plt

```

```{r}

## pull out cells in specific grid regions
int <- sf::st_intersection(seq, grid[c(22,23,28,29)])
cells2 <- seq[rownames(int),]

## grid 11 real
plt <- crawdad::vizAllClusters(cells = cells2,
                               coms = cells2$celltypes,
                               ofInterest = c("Lateral plate mesoderm", "Intermediate mesoderm"),
                               title = "grid 22,23,28,29",
                               axisAdj = 1, s = 10, a = 0.5) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  ggplot2::theme(legend.position="none")

ggplot2::ggsave(filename = "2A_mesoderm_tissue_zoom.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))

plt

```

## Fig 2 - mesoderms trends

```{r}

grDevices::pdf(file = paste0(figpath, "/2B_mesoderm_trends.pdf"), width = 4, height = 4)
dat <- crawdad::meltResultsList(results, id = "dist100")
d <- dat[dat$reference == "Lateral plate mesoderm" & dat$neighbor == "Intermediate mesoderm",]
crawdad::plotTrends(results = d, idcol = "id")
dev.off()

# dat <- crawdad::meltResultsList(results, id = "dist100")
# d <- dat[dat$reference == "Intermediate mesoderm" & dat$neighbor == "Lateral plate mesoderm",]
# crawdad::plotTrends(results = d, idcol = "id")
# 
# d <- dat[dat$reference == "Intermediate mesoderm",]
# crawdad::plotTrendsOverlay(results = d)
# 
# d <- dat[dat$reference == "Lateral plate mesoderm",]
# crawdad::plotTrendsOverlay(results = d)

```

```{r}

dat <- crawdad::meltResultsList(results)

d1 <- dat[dat$reference == "Lateral plate mesoderm" & dat$neighbor == "Intermediate mesoderm",]
plt <- vizTrends(dat = d1) +
  ggplot2::scale_x_log10()
plt 

d2 <- dat[dat$reference == "Endothelium" & dat$neighbor == "Haematoendothelial progenitors",]
plt <- vizTrends(dat = d2) +
  ggplot2::scale_x_log10()
plt 

## combine the trends into one data.frame, and have the "id" column label the combo, plot both lines on same plot
## turn off the facet wrap so just coloring the two trends, which are labeled using the "id" column

d1$id <- "Lateral plate mesoderm vs Intermediate mesoderm"
d2$id <- "Endothelium vs Haematoendothelial progenitors"

d <- dplyr::bind_rows(list(d1, d2))

plt <- vizTrends(dat = d, facet = FALSE) +
  ggplot2::scale_x_log10() +
  ggplot2::theme(legend.position="none")
plt 

ggplot2::ggsave(filename = "2B_mesoderm_endo_trends.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 4,
                height = 4,
                units = c("in"))


```


## Fig 2 - endothelium on tissue

Endothelium
Haematoendothelial progenitors

```{r, fig.height=4, fig.width=4}

plt <- crawdad::vizAllClusters(cells = seq,
                        coms = as.factor(seq$celltypes),
                        ofInterest = c("Endothelium", "Haematoendothelial progenitors"),
                        s = 2) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))
  ggplot2::theme(legend.position="none")

ggplot2::ggsave(filename = "2A_endothelium_tissue.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))
plt

```

### zoom in

```{r}

## shuffling grid
grid <- sf::st_make_grid(seq, cellsize = 1000)
grid_coords <- as.data.frame(sf::st_coordinates(grid))

## centers of the grids to add the tile IDs
grid_coords_centroids <- as.data.frame(sf::st_coordinates(sf::st_centroid(grid)))
grid_coords_centroids$name <- as.character(rownames(grid_coords_centroids))


plt <- crawdad::vizAllClusters(cells = seq,
                               coms = seq$celltypes,
                               ofInterest = c("Endothelium", "Haematoendothelial progenitors"),
                               title = "seq",
                               axisAdj = 1, s = 6, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y") +
  
  ## add in the grid information on top of the plot
  ggplot2::geom_sf(data = grid, fill = NA) +
  ggplot2::geom_text(data = grid_coords_centroids, ggplot2::aes(X, Y, label = name))
  
plt

```

```{r}

## pull out cells in specific grid regions
int <- sf::st_intersection(seq, grid[c(22,23,28,29)])
cells2 <- seq[rownames(int),]

## grid 11 real
plt <- crawdad::vizAllClusters(cells = cells2,
                               coms = cells2$celltypes,
                               ofInterest = c("Endothelium", "Haematoendothelial progenitors"),
                               title = "grid 22,23,28,29",
                               axisAdj = 1, s = 10, a = 0.5) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  ggplot2::theme(legend.position="none")
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")

ggplot2::ggsave(filename = "2A_endothelium_tissue_zoom.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))

plt

```

## Fig 2 - endothelium trends

```{r}

grDevices::pdf(file = paste0(figpath, "/2B_endothelium_trends.pdf"), width = 4, height = 4)
dat <- crawdad::meltResultsList(results, id = "dist100")
d <- dat[dat$reference == "Endothelium" & dat$neighbor == "Haematoendothelial progenitors",]
crawdad::plotTrends(results = d, idcol = "id")
dev.off()

dat <- crawdad::meltResultsList(results, id = "dist100")
d <- dat[dat$reference == "Haematoendothelial progenitors" & dat$neighbor == "Endothelium",]
crawdad::plotTrends(results = d, idcol = "id")

d <- dat[dat$reference == "Haematoendothelial progenitors",]
crawdad::plotTrendsOverlay(results = d)

d <- dat[dat$reference == "Endothelium",]
crawdad::plotTrendsOverlay(results = d)

```






