---
title: "seqFISH Analysis"
author: "Brendan F. Miller"
date: "2/15/2023"
output: rmarkdown::html_document
# output:
#   md_document:
#     variant: markdown_github
vignette: >
  %\VignetteIndexEntry{seqFISH analysis}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

This vignette will go through analyses to reproduce the results and figures of the SeqFISH dataset.

```{r}

library(crawdad)
library(dplyr)

```

```{r}

ncores = 2

```

# Load data

```{r, eval=FALSE}

seq <- read.csv2(file = paste0(here::here(), "/data/seqfish/seqfish.meta.csv"), row.names = 1, sep = ",")
seq <- seq[,c("x", "y", "cluster")]
## make sure the coordinates are numeric
seq <- seq %>%
  dplyr::mutate_at(vars(x, y), as.numeric)

## convert to SP
seq <- crawdad:::toSP(pos = seq[,c("x", "y")],
                        celltypes = seq$cluster)
seq

```

```{r}
## the above dataset has been saved as an rda file and able to load here:
data(seq)

## convert to SP
seq <- crawdad:::toSP(pos = seq[,c("x", "y")],
                        celltypes = seq$celltypes)
seq

```

# Visualize celltypes

```{r, fig.height=24, fig.width=24}

crawdad::vizEachCluster(cells = seq,
                        coms = as.factor(seq$celltypes),
                        s = 2)

```

# Make shuffled background

```{r}

scales = c(100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1200, 1500, 3000, 6000)

```

```{r, eval=FALSE}

oldw <- getOption("warn")
options(warn = -1)

## generate background
shuffle.list <- crawdad:::makeShuffledCells(seq,
                          scales = scales,
                          perms = 1,
                          ncores = ncores,
                          seed = 1,
                          verbose = TRUE)

options(warn = oldw)

## note: 0.77 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7
```

save shuffled object
```{r, eval=FALSE}

saveRDS(shuffle.list, paste0(here::here(), "/data/seqfish/sp.seqfish.shuffled_res100-6000.rds"))

```

# Run pairwise analysis

```{r, eval=FALSE}

oldw <- getOption("warn")
options(warn = -1)

## find trends, passing background as parameter
results <- crawdad::findTrends(seq,
                        dist = 100,
                        shuffle.list = shuffle.list,
                        ncores = ncores,
                        verbose = TRUE)

options(warn = oldw)

## note: 1.02 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7

```

save pairwise results object
```{r, eval=FALSE}

saveRDS(results, paste0(here::here(), "/data/seqfish/sp.seqfish.pairwise.results.dist100.res100-6000.rds"))

## object with distances 50-300: sp.seqfish.pairwise.50-300.results.res100-6000.removeDups.rds

```

# Filter trends

```{r}

results <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.pairwise.results.dist100.res100-6000.rds"))

results.coloc <- crawdad::filterCoTrends(results = results, alpha = 0.05)
results.sep <- crawdad::filterSepTrends(results = results, alpha = 0.05)
results.change <- crawdad::filterChangeTrends(results = results, alpha = 0.05)

```

# Defining subsets

```{r, eval=FALSE}

oldw <- getOption("warn")
options(warn = -1)

binomMat <- crawdad::binomialTestMatrix(seq,
                               neigh.dist = 100,
                               ncores = ncores,
                               verbose = TRUE)

head(binomMat)

options(warn = oldw)

## note: 2.49 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7
## previous code iteration on rockfish took 21 minutes with 32 cores! much better now 

```

save the binomial matrix
```{r, eval=FALSE}

saveRDS(binomMat, paste0(here::here(), "/data/seqfish/sp.seqfish.binomMat.near.subdist100.rds"))

```

```{r, eval=FALSE}

oldw <- getOption("warn")
options(warn = -1)

subset.list <- crawdad::selectSubsets(binomMat,
                             seq$celltypes,
                             sub.type = "near",
                             sub.thresh = 0.05,
                             ncores = ncores,
                             verbose = TRUE)

options(warn = oldw)

```

save the subsets
```{r, eval=FALSE}

saveRDS(subset.list, paste0(here::here(), "/data/seqfish/sp.seqfish.subsets.near.subdist100.rds"))

```

# Run analysis on subsets

```{r, eval=FALSE}

oldw <- getOption("warn")
options(warn = -1)

results.subsets <- crawdad::findTrends(seq,
                        dist = 100,
                        shuffle.list = shuffle.list,
                        subset.list = subset.list,
                        ncores = ncores,
                        verbose = TRUE)

options(warn = oldw)

## note: 11.68 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7; vs 7.53 minutes 14 cores on rockfish

```

save the results
```{r, eval=FALSE}

saveRDS(results.subsets, paste0(here::here(), "/data/seqfish/sp.seqfish.subset.results.dist100.res100-6000.rds"))

## sp.seqfish.triplet.near.binom.subdist100.dist100.results.res100-6000.removeDups.rds

```

# ---------------------------------
# Figures

stored on `brendan` branch in:
`/plots/seqfish/`

## load data

```{r}

## folder on `brendan` branch
figpath <- paste0(here::here(), "/plots/seqfish")

shuffle.list <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.shuffled_res100-6000.rds"))

binomMat <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.binomMat.near.subdist100.rds"))
subset.list <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.subsets.near.subdist100.rds"))

results <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.pairwise.results.dist100.res100-6000.rds"))
results.subsets <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.subset.results.dist100.res100-6000.rds"))

shuffle.list2 <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.shuffled_perms_res100-6000.rds"))
results2 <- readRDS(paste0(here::here(), "/data/seqfish/sp.seqfish.results_perms10.dist100.res100-6000.rds"))

```

## Fig 2 - mesoderms on tissue

```{r, fig.height=4, fig.width=4}

plt <- crawdad::vizAllClusters(cells = seq,
                        coms = as.factor(seq$celltypes),
                        ofInterest = c("Lateral plate mesoderm", "Intermediate mesoderm"),
                        s = 2) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))
  ggplot2::theme(legend.position="none")
plt

# ggplot2::ggsave(filename = "2_seqfish_mesoderm_tissue.pdf",
#                 plot = plt,
#                 device = "pdf",
#                 path = figpath,
#                 scale = 1,
#                 width = 6,
#                 height = 6,
#                 units = c("in"))

```

```{r, eval=FALSE}

## shuffling grid
grid <- sf::st_make_grid(seq, cellsize = 1000)

## centers of the grids to add the tile IDs
grid_coords_centroids <- as.data.frame(sf::st_coordinates(sf::st_centroid(grid)))
grid_coords_centroids$name <- as.character(rownames(grid_coords_centroids))


plt <- crawdad::vizAllClusters(cells = seq,
                        coms = as.factor(seq$celltypes),
                        ofInterest = c("Lateral plate mesoderm", "Intermediate mesoderm"),
                        s = 2) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))
  ggplot2::theme(legend.position="none") +
  # ggplot2::theme(legend.position="none") +
  
  ## add in the grid information on top of the plot
  ggplot2::geom_sf(data = grid, fill = NA) 
  # ggplot2::geom_text(data = grid_coords_centroids, ggplot2::aes(X, Y, label = name))
plt

ggplot2::ggsave(filename = "2B_mesoderm_tissue_grid.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))

```

```{r, eval=FALSE}

## pull out cells in specific grid regions
int <- sf::st_intersection(seq, grid[c(22,23,28,29)])
cells2 <- seq[rownames(int),]

## grid 11 real
plt <- crawdad::vizAllClusters(cells = cells2,
                               coms = cells2$celltypes,
                               ofInterest = c("Lateral plate mesoderm", "Intermediate mesoderm"),
                               title = "grid 22,23,28,29",
                               axisAdj = 1, s = 10, a = 0.5) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  ggplot2::theme(legend.position="none")
plt

ggplot2::ggsave(filename = "2_seqfish_mesoderm_tissue_zoom.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))

```

## Fig 2 - endothelium on tissue

Endothelium
Haematoendothelial progenitors

```{r, fig.height=4, fig.width=4}

plt <- crawdad::vizAllClusters(cells = seq,
                        coms = as.factor(seq$celltypes),
                        ofInterest = c("Endothelium", "Haematoendothelial progenitors"),
                        s = 2) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))
  ggplot2::theme(legend.position="none")
plt

# ggplot2::ggsave(filename = "2_seqfish_endothelium_tissue.pdf",
#                 plot = plt,
#                 device = "pdf",
#                 path = figpath,
#                 scale = 1,
#                 width = 6,
#                 height = 6,
#                 units = c("in"))

```

```{r, eval=FALSE}

## shuffling grid
grid <- sf::st_make_grid(seq, cellsize = 1000)

## centers of the grids to add the tile IDs
grid_coords_centroids <- as.data.frame(sf::st_coordinates(sf::st_centroid(grid)))
grid_coords_centroids$name <- as.character(rownames(grid_coords_centroids))


plt <- crawdad::vizAllClusters(cells = seq,
                        coms = as.factor(seq$celltypes),
                        ofInterest = c("Endothelium", "Haematoendothelial progenitors"),
                        s = 2) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))
  ggplot2::theme(legend.position="none") +
  
  ## add in the grid information on top of the plot
  ggplot2::geom_sf(data = grid, fill = NA) +
  ggplot2::geom_text(data = grid_coords_centroids, ggplot2::aes(X, Y, label = name))
plt

ggplot2::ggsave(filename = "2B_endothelium_tissue_grid.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))

```

```{r, eval=FALSE}

## pull out cells in specific grid regions
int <- sf::st_intersection(seq, grid[c(22,23,28,29)])
cells2 <- seq[rownames(int),]

## grid 11 real
plt <- crawdad::vizAllClusters(cells = cells2,
                               coms = cells2$celltypes,
                               ofInterest = c("Endothelium", "Haematoendothelial progenitors"),
                               title = "grid 22,23,28,29",
                               axisAdj = 1, s = 10, a = 0.5) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  ggplot2::theme(legend.position="none")
  # ggplot2::theme(legend.position="none") +
plt

ggplot2::ggsave(filename = "2_seqfish_endothelium_tissue_zoom.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))

```

## Fig 2 - trends

```{r}

dat <- crawdad::meltResultsList(results)
d <- dat[dat$reference %in% c("Lateral plate mesoderm", "Intermediate mesoderm", "Endothelium", "Haematoendothelial progenitors") & dat$neighbor %in% c("Lateral plate mesoderm", "Intermediate mesoderm", "Endothelium", "Haematoendothelial progenitors"),]

plt <- vizTrends.heatmap(dat = d)
plt

# ggplot2::ggsave(filename = "2_slideseq_trends_heatmap.pdf",
#                 plot = plt,
#                 device = "pdf",
#                 path = figpath,
#                 scale = 1,
#                 width = 9,
#                 height = 6,
#                 units = c("in"))

```

## Fig 2 - trends with permutations

```{r, eval=FALSE}

data(seq)

## convert to SP
seq <- crawdad:::toSP(pos = seq[,c("x", "y")],
                        celltypes = seq$celltypes)

ncores = 2
scales = c(100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1200, 1500, 3000, 6000)

oldw <- getOption("warn")
options(warn = -1)

## generate background
shuffle.list2 <- crawdad:::makeShuffledCells(seq,
                          scales = scales,
                          perms = 10,
                          ncores = ncores,
                          verbose = TRUE)

options(warn = oldw)

```

```{r, eval=FALSE}

saveRDS(shuffle.list2, paste0(here::here(), "/data/seqfish/sp.seqfish.shuffled_perms_res100-6000.rds"))

```

```{r, eval=FALSE}

oldw <- getOption("warn")
options(warn = -1)

## find trends, passing background as parameter
results2 <- crawdad::findTrends(seq,
                        dist = 100,
                        shuffle.list = shuffle.list2,
                        ncores = ncores,
                        verbose = TRUE,
                        returnMeans = FALSE)

options(warn = oldw)

```

```{r, eval=FALSE}

saveRDS(results2, paste0(here::here(), "/data/seqfish/sp.seqfish.results_perms10.dist100.res100-6000.rds"))

```

```{r}

dat <- crawdad::meltResultsList(results2, withPerms = TRUE)
d1 <- dat[dat$reference == "Lateral plate mesoderm" & dat$neighbor == "Intermediate mesoderm",]
d2 <- dat[dat$reference == "Endothelium" & dat$neighbor == "Haematoendothelial progenitors",]
## combine the trends into one data.frame, and have the "id" column label the combo, plot both lines on same plot
## turn off the facet wrap so just coloring the two trends, which are labeled using the "id" column
d1$id <- "Lateral plate mesoderm with Intermediate mesoderm"
d2$id <- "Endothelium with Haematoendothelial progenitors"
d <- dplyr::bind_rows(list(d1, d2))
# scales as numeric:
d <- d %>%
  dplyr::mutate_at(vars(scale), as.numeric)

# mean_line <- d %>%
#   group_by(reference, scale) %>%
#   summarise(Mean = mean(Z))

plt <- vizTrends(dat = d1, facet = TRUE, lines = FALSE) +
  ggplot2::geom_smooth(data = d1, 
                       ggplot2::aes(x = scale, y = Z, color = id, fill = id)) +
  ggplot2::scale_x_log10() +
  ggplot2::ylim(c(-3, 15)) +
  ggplot2::theme(legend.position="none")
plt

plt <- vizTrends(dat = d2, facet = TRUE, lines = FALSE) +
  ggplot2::geom_smooth(data = d2, 
                       ggplot2::aes(x = scale, y = Z, color = id, fill = id)) +
  ggplot2::scale_x_log10() +
  ggplot2::ylim(c(-3, 15)) +
  ggplot2::theme(legend.position="none")
plt

# ggplot2::ggsave(filename = "2_seqfish_trends_permutations_endo.pdf",
#                 plot = plt,
#                 device = "pdf",
#                 path = figpath,
#                 scale = 1,
#                 width = 6,
#                 height = 6,
#                 units = c("in"))

```

