---
title: "Slide-seq"
author: "Brendan F. Miller"
date: "2/15/2023"
output: rmarkdown::html_document
# output:
#   md_document:
#     variant: markdown_github
vignette: >
  %\VignetteIndexEntry{Slide-seq analysis}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

This vignette will go through analyses to reproduce the results and figures of the SeqFISH dataset.

```{r}

library(crawdad)
library(dplyr)

```

```{r}

ncores = 7

```

# Load data

```{r, eval=FALSE}

slide <- read.csv2(file = paste0(here::here(), "/data/slideseq/slideseq.puck190926_08.rctd.meta.csv.gz"), row.names = 1, sep = ",")
slide <- slide[,c("x", "y", "class")]
## make sure the coordinates are numeric
slide <- slide %>%
  dplyr::mutate_at(vars(x, y), as.numeric)

## convert to SP
slide <- crawdad:::toSP(pos = slide[,c("x", "y")],
                        celltypes = slide$class)
slide

```

```{r}
## the above dataset has been saved as an rda file and able to load here:
data(slide)

## convert to SP
slide <- crawdad:::toSP(pos = slide[,c("x", "y")],
                        celltypes = slide$celltypes)
slide

```

# Visualize celltypes

```{r, fig.height=24, fig.width=24}

crawdad::vizEachCluster(cells = slide,
                        coms = as.factor(slide$celltypes),
                        s = 2)

```

# Make shuffled background

```{r}

resolutions <- c(30, 50, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1200, 1500, 3000, 6000)

```

```{r}
## generate background
shuffle.list <- crawdad:::makeShuffledCells(slide,
                          resolutions = resolutions,
                          perms = 1,
                          ncores = ncores,
                          seed = 1,
                          verbose = TRUE)

## note: 0.34 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7
```

save shuffled object
```{r, eval=FALSE}

saveRDS(shuffle.list, paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.shuffled_res30-6000.rds"))

## slideseqPuck.190926_08.rctd.shuffled_res30-6000.rds

```

# Run pairwise analysis

```{r}

## find trends, passing background as parameter
results <- crawdad::findTrends(slide,
                        dist = 100,
                        shuffle.list = shuffle.list,
                        perms = 1,
                        ncores = ncores,
                        seed = 1, 
                        verbose = TRUE)

## note: 1.3 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7

```

save pairwise results object
```{r, eval=FALSE}

saveRDS(results, paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.pairwise.results.dist100.res30-6000.rds"))

## object with distances 30-300: slideseqPuck.190926_08.rctd.pairwise.30-300.results.removeDups.rds

```

# Visualizing trends

```{r}

crawdad::plotTrends(results = results["Purkinje"], idcol = "dist")

```

```{r}

crawdad::plotTrends(results = results["Bergmann"], idcol = "dist")

```

# Defining susbets

```{r}

binomMat <- crawdad::binomialTestMatrix(slide,
                               neigh.dist = 100,
                               ncores = ncores,
                               verbose = TRUE)

head(binomMat)

## note: 0.8 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7
## previous code iteration on rockfish took 3:56 minutes with 32 cores! much better now 

```

save the binomial matrix
```{r, eval=FALSE}

saveRDS(binomMat, paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.binomMat.near.subdist100.rds"))

```

```{r}

subset.list <- crawdad::selectSubsets(binomMat,
                             slide$celltypes,
                             sub.type = "near",
                             sub.thresh = 0.05,
                             ncores = ncores,
                             verbose = TRUE)

```

save the subsets
```{r, eval=FALSE}

saveRDS(subset.list, paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.subsets.near.subdist100.rds"))

## slideseqPuck.190926_08.rctd.subsets.near.subdist100.rds

```

# Run analysis on subsets

```{r}

results.subsets <- crawdad::findTrends(slide,
                        dist = 100,
                        shuffle.list = shuffle.list,
                        subset.list = subset.list,
                        perms = 1,
                        ncores = ncores,
                        seed = 1, 
                        verbose = TRUE)

## note: 12.49 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7; vs 2:46 minutes 14 cores on rockfish

```

save the results
```{r, eval=FALSE}

saveRDS(results.subsets, paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.subset.results.dist100.res30-6000.rds"))

## slideseqPuck.190926_08.rctd.near.binom.subdist100.dist100.results.removeDups.rds

```

# ---------------------------------
# Figures

## load data

```{r, eval=FALSE}

shuffle.list <- readRDS(paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.shuffled_res30-6000.rds"))

binomMat <- readRDS(paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.binomMat.near.subdist100.rds"))
subset.list <- readRDS(paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.subsets.near.subdist100.rds"))

results <- readRDS(paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.pairwise.results.dist100.res30-6000.rds"))
results.subsets <- readRDS(paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.subset.results.dist100.res30-6000.rds"))

```




