---
title: "Slide-seq"
author: "Brendan F. Miller"
date: "2/15/2023"
output: rmarkdown::html_document
# output:
#   md_document:
#     variant: markdown_github
vignette: >
  %\VignetteIndexEntry{Slide-seq analysis}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

This vignette will go through analyses to reproduce the results and figures of the Slide-seq dataset.

```{r}

library(crawdad)
library(dplyr)

```

```{r}

ncores = 7

```

# Load data

```{r, eval=FALSE}

slide <- read.csv2(file = paste0(here::here(), "/data/slideseq/slideseq.puck190926_08.rctd.meta.csv.gz"), row.names = 1, sep = ",")
slide <- slide[,c("x", "y", "class")]
## make sure the coordinates are numeric
slide <- slide %>%
  dplyr::mutate_at(vars(x, y), as.numeric)

## convert to SP
slide <- crawdad:::toSP(pos = slide[,c("x", "y")],
                        celltypes = slide$class)
slide

```

```{r}
## the above dataset has been saved as an rda file and able to load here:
data(slide)

## convert to SP
slide <- crawdad:::toSP(pos = slide[,c("x", "y")],
                        celltypes = slide$celltypes)
slide

```

# Visualize celltypes

```{r, fig.height=24, fig.width=24}

crawdad::vizEachCluster(cells = slide,
                        coms = as.factor(slide$celltypes),
                        s = 2)

```

# Make shuffled background

```{r}

resolutions <- c(30, 50, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1200, 1500, 3000, 6000)

```

```{r}
## generate background
shuffle.list <- crawdad:::makeShuffledCells(slide,
                          resolutions = resolutions,
                          perms = 1,
                          ncores = ncores,
                          seed = 1,
                          verbose = TRUE)

## note: 0.34 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7
```

save shuffled object
```{r, eval=FALSE}

saveRDS(shuffle.list, paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.shuffled_res30-6000.rds"))

## slideseqPuck.190926_08.rctd.shuffled_res30-6000.rds

```

# Run pairwise analysis

```{r}

## find trends, passing background as parameter
results <- crawdad::findTrends(slide,
                        dist = 100,
                        shuffle.list = shuffle.list,
                        ncores = ncores,
                        verbose = TRUE)

## note: 1.3 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7

```

save pairwise results object
```{r, eval=FALSE}

saveRDS(results, paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.pairwise.results.dist100.res30-6000.rds"))

## object with distances 30-300: slideseqPuck.190926_08.rctd.pairwise.30-300.results.removeDups.rds

```

# Visualizing trends

```{r}

crawdad::plotTrends(results = results["Purkinje"], idcol = "dist")

```

```{r}

crawdad::plotTrends(results = results["Bergmann"], idcol = "dist")

```

# Defining susbets

```{r}

binomMat <- crawdad::binomialTestMatrix(slide,
                               neigh.dist = 100,
                               ncores = ncores,
                               verbose = TRUE)

head(binomMat)

## note: 0.8 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7
## previous code iteration on rockfish took 3:56 minutes with 32 cores! much better now 

```

save the binomial matrix
```{r, eval=FALSE}

saveRDS(binomMat, paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.binomMat.near.subdist100.rds"))

```

```{r}

subset.list <- crawdad::selectSubsets(binomMat,
                             slide$celltypes,
                             sub.type = "near",
                             sub.thresh = 0.05,
                             ncores = ncores,
                             verbose = TRUE)

```

save the subsets
```{r, eval=FALSE}

saveRDS(subset.list, paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.subsets.near.subdist100.rds"))

## slideseqPuck.190926_08.rctd.subsets.near.subdist100.rds

```

# Run analysis on subsets

```{r}

results.subsets <- crawdad::findTrends(slide,
                        dist = 100,
                        shuffle.list = shuffle.list,
                        subset.list = subset.list,
                        ncores = ncores,
                        verbose = TRUE)

## note: 12.49 minutes with 7 cores, 2.3 GHz Quad-Core Intel Core i7; vs 2:46 minutes 14 cores on rockfish

```

save the results
```{r, eval=FALSE}

saveRDS(results.subsets, paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.subset.results.dist100.res30-6000.rds"))

## slideseqPuck.190926_08.rctd.near.binom.subdist100.dist100.results.removeDups.rds

```

# ---------------------------------
# Figures

## load data

```{r, eval=FALSE}

## folder on `brendan` branch
figpath <- paste0(here::here(), "/plots/slideseq")

shuffle.list <- readRDS(paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.shuffled_res30-6000.rds"))

binomMat <- readRDS(paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.binomMat.near.subdist100.rds"))
subset.list <- readRDS(paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.subsets.near.subdist100.rds"))

results <- readRDS(paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.pairwise.results.dist100.res30-6000.rds"))
results.subsets <- readRDS(paste0(here::here(), "/data/slideseq/slideseqPuck.190926_08.subset.results.dist100.res30-6000.rds"))

```

## Fig 3 - purk nad berg on tissue

```{r, fig.height=4, fig.width=4}

plt <- crawdad::vizAllClusters(cells = slide,
                        coms = as.factor(slide$celltypes),
                        ofInterest = c("Bergmann", "Purkinje"),
                        s = 2) +
  ggplot2::theme(legend.position="none")
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))

ggplot2::ggsave(filename = "3A_purk_berg_tissue.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))
plt

```

### zoom in

```{r}

## shuffling grid
grid <- sf::st_make_grid(slide, cellsize = 1000)
grid_coords <- as.data.frame(sf::st_coordinates(grid))

## centers of the grids to add the tile IDs
grid_coords_centroids <- as.data.frame(sf::st_coordinates(sf::st_centroid(grid)))
grid_coords_centroids$name <- as.character(rownames(grid_coords_centroids))


plt <- crawdad::vizAllClusters(cells = slide,
                               coms = slide$celltypes,
                               ofInterest = c("Bergmann", "Purkinje"),
                               title = "seq",
                               axisAdj = 1, s = 6, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y") +
  
  ## add in the grid information on top of the plot
  ggplot2::geom_sf(data = grid, fill = NA) +
  ggplot2::geom_text(data = grid_coords_centroids, ggplot2::aes(X, Y, label = name))
  
plt

```

```{r}

## pull out cells in specific grid regions
int <- sf::st_intersection(slide, grid[[14]])
cells2 <- slide[rownames(int),]

## grid 11 real
plt <- crawdad::vizAllClusters(cells = cells2,
                               coms = cells2$celltypes,
                               ofInterest = c("Bergmann", "Purkinje"),
                               title = "grid14",
                               axisAdj = 1, s = 10, a = 0.5) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")

ggplot2::ggsave(filename = "3A_purk_berg_tissue_zoom.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))

plt

```

## Fig 3 - purk and berg trends

```{r}

grDevices::pdf(file = paste0(figpath, "/3B_berg_trends.pdf"), width = 4, height = 4)
dat <- crawdad::meltResultsList(results, id = "dist100")
d <- dat[dat$reference == "Bergmann" & dat$neighbor == "Purkinje",]
crawdad::plotTrends(results = d, idcol = "id")
dev.off()

grDevices::pdf(file = paste0(figpath, "/3B_purk_trends.pdf"), width = 4, height = 4)
dat <- crawdad::meltResultsList(results, id = "dist100")
d <- dat[dat$reference == "Purkinje" & dat$neighbor == "Bergmann",]
crawdad::plotTrends(results = d, idcol = "id")
dev.off()


d <- dat[dat$reference == "Purkinje",]
crawdad::plotTrendsOverlay(results = d)

d <- dat[dat$reference == "Bergmann",]
crawdad::plotTrendsOverlay(results = d)

```

```{r}

dat <- crawdad::meltResultsList(results)

d1 <- dat[dat$reference == "Purkinje" & dat$neighbor == "Bergmann",]
plt <- vizTrends(dat = d1) +
  ggplot2::scale_x_log10()
plt 

d2 <- dat[dat$reference == "Bergmann" & dat$neighbor == "Purkinje",]
plt <- vizTrends(dat = d2) +
  ggplot2::scale_x_log10()
plt 

## combine the trends into one data.frame, and have the "id" column label the combo, plot both lines on same plot
## turn off the facet wrap so just coloring the two trends, which are labeled using the "id" column

d1$id <- "Purkinje vs Bergmann"
d2$id <- "Bergmann vs Purkinje"

d <- dplyr::bind_rows(list(d1, d2))

plt <- vizTrends(dat = d, facet = FALSE) +
  ggplot2::scale_x_log10() +
  ggplot2::theme(legend.position="none")
plt 

ggplot2::ggsave(filename = "3B_purkinje_bergmann_trends.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 4,
                height = 4,
                units = c("in"))


```

## Fig 3 - ubc on tissue

```{r, fig.height=4, fig.width=4}

plt <- crawdad::vizAllClusters(cells = slide,
                        coms = as.factor(slide$celltypes),
                        ofInterest = c("UBCs", "Granule"),
                        s = 2) +
  ggplot2::theme(legend.position="none")
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1))

ggplot2::ggsave(filename = "3A_UBCs_tissue.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))
plt

```

### zoom in

```{r}

## shuffling grid
grid <- sf::st_make_grid(slide, cellsize = 1000)
grid_coords <- as.data.frame(sf::st_coordinates(grid))

## centers of the grids to add the tile IDs
grid_coords_centroids <- as.data.frame(sf::st_coordinates(sf::st_centroid(grid)))
grid_coords_centroids$name <- as.character(rownames(grid_coords_centroids))


plt <- crawdad::vizAllClusters(cells = slide,
                               coms = slide$celltypes,
                               ofInterest = c("UBCs", "Granule"),
                               title = "seq",
                               axisAdj = 1, s = 6, a = 0.5) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  # ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y") +
  
  ## add in the grid information on top of the plot
  ggplot2::geom_sf(data = grid, fill = NA) +
  ggplot2::geom_text(data = grid_coords_centroids, ggplot2::aes(X, Y, label = name))
  
plt

```

```{r}

## pull out cells in specific grid regions
int <- sf::st_intersection(slide, grid[[14]])
cells2 <- slide[rownames(int),]

## grid 11 real
plt <- crawdad::vizAllClusters(cells = cells2,
                               coms = cells2$celltypes,
                               ofInterest = c("UBCs", "Granule"),
                               title = "grid14",
                               axisAdj = 1, s = 10, a = 0.5) +
  # ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size=2), ncol = 1)) +
  ggplot2::theme(legend.position="none") +
  ggplot2::labs(x = "x",
                y = "y")

ggplot2::ggsave(filename = "3A_UBCs_tissue_zoom.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 6,
                height = 6,
                units = c("in"))

plt

```

## Fig 3 - ubc trends

```{r}

grDevices::pdf(file = paste0(figpath, "/3B_ubc_trends.pdf"), width = 4, height = 4)
dat <- crawdad::meltResultsList(results, id = "dist100")
d <- dat[dat$reference == "UBCs" & dat$neighbor == "Granule",]
crawdad::plotTrends(results = d, idcol = "id")
dev.off()

grDevices::pdf(file = paste0(figpath, "/3B_granule_trends.pdf"), width = 4, height = 4)
dat <- crawdad::meltResultsList(results, id = "dist100")
d <- dat[dat$reference == "Granule" & dat$neighbor == "UBCs",]
crawdad::plotTrends(results = d, idcol = "id")
dev.off()


d <- dat[dat$reference == "UBCs",]
crawdad::plotTrendsOverlay(results = d)

d <- dat[dat$reference == "Granule",]
crawdad::plotTrendsOverlay(results = d)

```

```{r}

dat <- crawdad::meltResultsList(results)

d1 <- dat[dat$reference == "UBCs" & dat$neighbor == "Granule",]
plt <- vizTrends(dat = d1) +
  ggplot2::scale_x_log10()
plt 

d2 <- dat[dat$reference == "Granule" & dat$neighbor == "UBCs",]
plt <- vizTrends(dat = d2) +
  ggplot2::scale_x_log10()
plt 

## combine the trends into one data.frame, and have the "id" column label the combo, plot both lines on same plot
## turn off the facet wrap so just coloring the two trends, which are labeled using the "id" column

d1$id <- "UBCs vs Granule"
d2$id <- "Granule vs UBCs"

d <- dplyr::bind_rows(list(d1, d2))

plt <- vizTrends(dat = d, facet = FALSE) +
  ggplot2::scale_x_log10() +
  ggplot2::theme(legend.position="none")
plt 

ggplot2::ggsave(filename = "3B_ubc_granule_trends.pdf",
                plot = plt,
                device = "pdf",
                path = figpath,
                scale = 1,
                width = 4,
                height = 4,
                units = c("in"))


```
